Contracts for Contracts

### 2022-12-13 paper.tex
intro

%% Title information
\title[Short Title]{Contracts for Contracts}         %% [Short Title] is optional;
%\title[Short Title]{Contracts meet Contracts}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Adding Higher-Order Contracts for Smart Contract}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

@@

reliability and security of smart contract programs

contracts enforce properties/check invariants at runtime
  Effiel, Design by Contract
  %https://se.inf.ethz.ch/~meyer/publications/old/dbc_chapter.pdf
  Racket, higher-order contracts

%https://docs.soliditylang.org/en/v0.8.17/control-structures.html?highlight=require#panic-via-assert-and-error-via-require
smart contracts use callback, need for higher-order contract,

One of the most adopted smart contract language Solidity indeed provides
linguistic mechanisms such as @require@ and @assert@ to enforce first-order
properties, however they does not deal with function values.
\todo{need double check}

Static verification cannot completely address the issue here since
transaction developers are oblivious to the user-provided callback.

It is unfortunate or even surprising that smart contract languages have not
adopted full-blown software contracts, especially higher-order contracts.

\subsubsection*{Contributions} This paper makes the following contributions:
\begin{itemize}
  \item We propose to add higher-order contracts to smart contract languages to
    improve the safety and reliability. We use Solidity to demonstrate our
    proposal.
  \item We present the core language ``Contract Solidity'' with higher-order
    contracts, including its syntax, formal semantics, and soundness.
  \item We examine a number of real-world cases and show that our proposal
    is effective in improving safety and reliability.
  \item We implement our approach on top of the existing Solidity compiler and shows
    that its runtime overhead is marginal.
\end{itemize}

@@

perhaps also performance evaluation?

@@

the implemention that modidifies the off-the-shelf Solidity compiler that
supports contracts and generates code with correct contracts checking

>>>>

%% Title information
\title[Short Title]{Contracts for Contracts}         %% [Short Title] is optional;
%\title[Short Title]{Contracts meet Contracts}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Adding Higher-Order Contracts for Smart Contracts}                     %% \subtitle is optional
%\subtitle{Expressive Assertion Monitoring System for Safer Smart Contracts}                     %% \subtitle is optional
\subtitle{Expressive Temporal Higher-Order Behavioral Contracts for Smart Contracts}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

%https://se.inf.ethz.ch/~meyer/publications/old/dbc_chapter.pdf
%https://docs.soliditylang.org/en/v0.8.17/control-structures.html?highlight=require#panic-via-assert-and-error-via-require

@@

%\todo{why? solidity is rather flexible, unsatisfying/weaker type system,
%not all invariants can be checked statically, e.g. the status of the chain}
%One \note{common} use case of higher-order functions in smart contracts is
%to express callback \todo{explain how it is used}.

%\todo{it is hard to enforce properties when callbacks exist}
%Static verification cannot completely address the issue here since
%transaction developers are oblivious to the user-provided callback.

%Building trustworthy and secure smart contract programs requires solid
%understanding and reliable enforcement of program invariants.

Smart contracts are programs to automate the execution of agreement so that all
parties involved in the agreement can be certain about the outcome.
Smart contracts are ``smart'' since by using blockchains they provide a
mechanism to \emph{ensure the economic outcome of transactions} without a
central authority.
A blockchain implements a consensus protocol of Byzantine-fault-tolerant
distributed ledgers.

It is however ironic that many popular smart contract languages, which hold the
great promise to ensure transactional outcomes via consensus and distributed
ledgers, do not provide enough means to \emph{ensure the quality of programs}.
Without ensuring the quality of programs, it is impossible to ensure
the economic outcome of transactions.

For example, Solidity provides first-order assertion primitives @require@ and
@assert@, but they cannot check function values (e.g. used as callbacks), which are a
common pattern used in smart contract programs. Attackers are able to utilize
the callback mechanism to \note{do crazy things and steal money}

The lack of sufficient linguistic means to specify and enforce subtle behaviors
discourages programmers to use higher-level abstractions when writing smart
contract code.
Instead, to be secure, programmers are obliged to write verbose code with
lower-level abstractions, which conflates the main business logic and
behavioral checks, leading to poor readability and maintainability.
For instance, to defend the famous DAO attack \note{cite} caused by allowing ``reentrancy''
via external functions, programmers have to be extra cautious about the execution
order or maintain additional states to check if a call is allowed.
Either way is not ideal.
And often worse, defensive checks are disregarded, leading to vulnerable code
causing real monetary lose.

Despite hindsight effort to audit, certify, or verify smart contract programs,
language designers are responsible to provide better languages so that
programmers are able to write invulnerable smart contracts in the first place.
In this paper, we argue that advanced \emph{behavioral contracts} provide a
lightweight linguistic solution for many security issues found in smart
contract programs.

\paragraph{\textbf{Behavioral Contracts for Smart Contracts}}
Behavioral contracts provide a mechanism for programmers to specify assumptions
and guarantees of software components. The specified assumptions and guarantees
will be monitored and reported when violated.
Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
Design-by-Contract methodology \cite{DBLP:conf/tools/Meyer98a}, behavioral
contracts have been widely adopted in Java, Haskell, C\#, Python, Racket, etc. \todo{cite}

\todo{why BC is good here?}

In this work, we design and implement a practical extension of Solidity, the
most popular language for building smart contract programs. The extension
enriches the Solidity language with \emph{temporal} and \emph{higher-order} behavioral
contracts, which checks \todo{...}
We develop a core model of the extension, including its syntax,
operational semantics, and discuss its soundness guarantee.

With the behavioral contract extension of Solidity, we examine
\note{N} cases of reported bugs and vulnerabilities and show that
our approach is effective in preventing these attacks.

\todo{discuss performance/gas consumption}

\subsubsection*{Contributions} This paper makes the following contributions:
\begin{itemize}
  \item We propose to add temporal and higher-order contracts to smart contract
    languages to improve the safety and reliability. We use Solidity to
    demonstrate our proposal.
  \item We present the core language ``Contract Solidity'' with our
    proposed extension, including its syntax, formal semantics, and soundness.
  \item We examine a number of real-world cases and show that our proposal is
    effective in improving safety and reliability.
  \item We implement our approach on top of the existing Solidity compiler and shows
    that its runtime overhead is marginal.
\end{itemize}

@@

\subsection{Reentrancy}

\subsection{Callback}

\todo{what else?}

@@

the implementation that modidifies the off-the-shelf Solidity compiler that
supports contracts and generates code with correct contracts checking

\section{Performance Evaluation}
